---
title: "DACA dashboard"
author: "Eduard Bukin"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
runtime: shiny
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
# library(ggwb)
library(flexdashboard)
library(shiny)
library(googleVis)

# Loading data
pd <-
  read_rds("clean-data/analysis_data.rds") %>%
  filter(!is.na(country_code)) %>%
  mutate(
    income = factor(income, levels = c("Low income", "Lower middle income", "Upper middle income", "High income")),
    country = factor(country, levels = sort(unique(pull(., country))))
  ) %>% 
  mutate(
    `Variable group` = ifelse(is.na(`Variable group`), "", `Variable group`),
    var = str_c(`Source`, " - ", ifelse(`Variable group` == '', '', str_c(`Variable group`, " - ")), var)) %>% 
  # select(-var_code, - Source, -`Variable group`, -flag)%>%
  # mutate(
  #   var = ifelse(Source == "GSMA", str_c(`Variable group`, " - ", var), var),
  #   var = ifelse(Source == "Digital Adoption Index", str_c("DAI - ", var), var),
  #   var = ifelse(Source == "WDI", str_c("WDI - ", var), var)
  # ) %>%
  arrange(Source, var, country, year)

bubles_data <- 
   read_rds("clean-data/google_vis_data.rds") 

axis_vars <-
  c(
    `NOTHING` = "nothing",
    set_names(
      distinct(pd, var, Source, var_code)$var_code,
      distinct(pd, var, var_code)$var
    )
  )

group_vars <-
  c(
    `Income group` = "income",
    `Region` = "region",
    `k-mean clusters` = "cluster"
  )

wb_col <- function(...) {
  wb_colors <- c(
    `Solid Blue` = "#002244",
    `Bright Blue` = "#009FDA",
    `orange` = "#F05023",
    `yellow` = "#FDB714",
    `red` = "#EB1C2D",
    `brown` = "#F78D28",
    `blue` = "#009CA7",
    `green` = "#00AB51",
    `purple` = "#872B90",
    `light_blue` = "#00A996"
  )

  cols <- c(...)

  if (is.null(cols)) {
    return(wb_colors)
  }

  wb_colors[cols]
}

wb_pal <- function(palette = "main", reverse = FALSE, ...) {
  wb_palettes <- list(
    `main` = wb_col("red", "yellow", "green", "blue", "purple")
  )

  pal <- wb_palettes[[palette]]

  if (reverse) pal <- rev(pal)

  colorRampPalette(pal, ...)
}
scale_color_wb <- function(palette = "main", discrete = TRUE, reverse = FALSE, ...) {
  pal <- wb_pal(palette = palette, reverse = reverse)

  if (discrete) {
    discrete_scale("colour", paste0("wb_", palette), palette = pal, ...)
  } else {
    scale_color_gradientn(colours = pal(256), ...)
  }
}

scale_fill_wb <- function(palette = "main", discrete = TRUE, reverse = FALSE, ...) {
  pal <- wb_pal(palette = palette, reverse = reverse)

  if (discrete) {
    discrete_scale("fill", paste0("wb_", palette), palette = pal, ...)
  } else {
    scale_fill_gradientn(colours = pal(256), ...)
  }
}


# Used functions
plot_scattered_countries <- function(plot_data, x_var, y_var, radius_var, colour_var, shape_var = NULL) {
  # if (radius_var == "nothing") radius_var <- NULL
  plot_data %>%
    drop_na() %>%
    ggplot() +
    aes_string(x = x_var, y = y_var, size = radius_var, colour = colour_var, fill = colour_var) +
    geom_point(position = "jitter", aes(alpha = 0.9)) +
    scale_color_wb(name = names(colour_var), discrete = TRUE) +
    scale_radius(name = names(radius_var)) +
    guides(
      fill = "none",
      alpha = "none",
      # shape = guide_legend(order = 2, override.aes = list(size=5, colour=wb_col("red"))),
      size = guide_legend(order = 1, override.aes = list(colour = wb_col("red"))),
      colour = guide_legend(order = 2, override.aes = list(size = 5))
    ) +
    xlab(names(x_var)) +
    ylab(names(y_var)) +
    theme_bw()
}
```



Agricultural transformation {data-orientation=rows}
=====================================  


Row {.tabset}
-------------------------------------

### Agricultural transformation 

```{r}
# create a ui interaction:
uiOutput("dynamic")

# render the ui:
output$dynamic <- renderUI({
  htmlOutput("data_bubles")
})

# here is your server activity:
output$data_bubles <- renderGvis({
  gvisMotionChart(
    data = bubles_data,
    idvar = "country",
    timevar = "year",
    xvar = "WDI - Employment in agriculture (% of total employment) (modeled ILO estimate)",
    yvar = "WDI - Agriculture, forestry, and fishing, value added (% of GDP)",
    colorvar = "region",
    sizevar = "Digital Adoption Index - Digital Adoption Index",
    date.format = "YYYY",
    options = list(width = "1200", height = "700")
  )
})

```


### Bubbles data download


```{r}
fluidPage(
  
fluidRow(
    DT::renderDataTable( #
      DT::datatable(
        bubles_data,
        rownames = FALSE,
        extensions = c("Buttons", "Scroller"),
        options = list(
          pageLength = 40,
          dom = c("Bfrtip"),
          scrollX = TRUE,
          # fixedColumns = list(leftColumns = 1),
          deferRender = TRUE,
          scrollY = 600,
          scroller = TRUE,
          buttons = list(
            list(extend = "copy", text = "Copy table"),
            list(extend = "excel", text = "Download table", title = "availability")
          )
        )
      )
    )
  )
)
```



Data and old agricultural transformation {data-orientation=rows}
=====================================  
    
Inputs {.sidebar data-width=400}
-------------------------------------


```{r}
selectInput("x_var", "X axis", axis_vars[axis_vars != "nothing"], axis_vars[axis_vars == "SL.AGR.EMPL.ZS"], FALSE, TRUE, width = "100%" )
selectInput("y_var", "Y axis", axis_vars[axis_vars != "nothing"], axis_vars[axis_vars == "NV.AGR.TOTL.ZS"], FALSE, TRUE, width = "100%")
```

You can set to have all bubles the same size. 
The bubles are all the same size by default if there is no data for the selected variable/year. 
To see what 

```{r}
checkboxInput("same_radius", "Same buble size", value = FALSE)
selectInput("radius_var", "SIze of the buble", axis_vars, axis_vars[axis_vars == "ndx.scr.gsm.ndx"], FALSE, selectize = TRUE, width = "100%")
sliderInput("active_year", "Year of analysis", 1990, 2017, 2015,
  step = 1, round = 0, ticks = TRUE,
  animate =
    animationOptions(
      interval = 1400,
      loop = FALSE,
      playButton = icon("play", lib = "glyphicon"),
      pauseButton = icon("pause", lib = "glyphicon")
    ),
  width = "100%",
  sep = ",",
  dragRange = FALSE
)
```

Here you specify variable for grouping countries into on colour group. If you select "cluster" k-means clustering based on the three selected variables (x, y and size) is performed.

```{r}
selectInput("colour_var", "Grouping variable", group_vars, group_vars[group_vars == "income"], FALSE, TRUE, width = "100%")
numericInput("number_of_clusters", "Provide number of clusters", 5, 2, 100, 1, width = "100%")
```



 
```{r}
colour_var <- reactive(input$colour_var)

filter_data <-
  reactive(
    {
      if (!input$same_radius) radius <- input$radius_var else radius <- NULL
      pd %>% 
        filter(var_code %in% c(input$x_var, input$y_var, radius)) %>%
        filter(year == input$active_year) 
    }
  )

export_data <-
  reactive({
    filter_data() %>%
      mutate(var = str_trunc(var, 25)) %>%
      select(country, region, income, year, val, flag, var) %>%
      {
        dta <- .
        ident <- dta %>% distinct(country, region, income, year)
        vals <- dta %>%
          select(country, year, val, var) %>%
          spread(var, val)
        flags <- dta %>%
          select(country, year, flag, var) %>%
          mutate(var = str_c(var, " - FLAG")) %>%
          spread(var, flag)
        ident %>%
          left_join(vals, c("country", "year")) %>%
          left_join(flags, c("country", "year")) %>%
          mutate_at(vars(contains("flag")), funs(ifelse(is.na(.), "Missing", .)))
      } %>% 
      arrange(country)
  })

zero_vars <-
  reactive({
    zero_vars <-
      filter_data() %>%
      group_by(var, var_code, flag) %>%
      count() %>%
      spread(flag, n) %>%
      ungroup() %>%
      mutate_all(funs(ifelse(is.na(.), 0, .))) %>%
      rename(Variable = var)

    if (!"Imputed" %in% names(zero_vars)) {
      zero_vars <-
        zero_vars %>%
        mutate(Imputed = 0)
    }
    zero_vars <- 
      zero_vars %>% 
      mutate(
        no_data = Original == 0 & Imputed == 0
      )
    zero_vars
  })

zero_message <-
  reactive({
    zero_vars() %>%
      as.list() %>%
      transpose() %>%
      map(.f = function(x) {
        tags$p(
          str_c(
            "Variable ", x$Variable, " contains: ",
            x$Original, " original, ",
            x$Imputed, " imputed and ",
            x$Missing, " missing values."
          )
        )
      }) %>%
      tagList() %>%
      return()
  })

plot_data <-
  reactive({
    contains_data <-
      zero_vars() %>%
      filter(!no_data) %>%
      pull(var_code)

    contains_zero <-
      zero_vars() %>%
      filter(no_data) %>%
      pull(var_code)

    if (!is.null(contains_data) & length(contains_data) < 2) return(NULL)

    flt_data <- filter_data()

    if (!is.null(contains_data) & length(contains_zero) > 0) {
      flt_data <-
        flt_data %>%
        filter(var_code != contains_zero)
    }

    flt_data %>%
      select(-var) %>%
      {
        dta <- .
        flags <-
          dta %>%
          group_by(country, year) %>%
          summarise(flag = ifelse(all(flag == "Original"), "Original",
            ifelse(all(flag == "Imputed"), "Imputed", "Partial")
          ))
        dta %>%
          select(-flag, -`Variable group`, -Source) %>%
          spread(var_code, val) %>%
          left_join(flags, by = c("country", "year"))
      } %>%
      mutate(
        flag = factor(flag, levels = c("Original", "Partial", "Imputed"))
      ) %>%
      drop_na()
  })

colours_calc <-
  reactive({
    if (colour_var() == "cluster") {
      radius <- input$radius_var
      if (is.na(radius)) radius <- NULL
      plot_data() %>%
        select(matches("\\.")) %>% 
        kmeans(input$number_of_clusters) %>%
        return()
    } else {
      return(NULL)
    }
  })

plot_data_2 <-
  reactive({
    req(plot_data())
    clusters <- colours_calc()

    if (colour_var() == "cluster" & !is.null(clusters)) {
      plot_data() %>%
        mutate(cluster = factor(clusters$cluster)) %>%
        return()
    } else {
      plot_data() %>%
        return()
    }
  })

build_plot <-
  reactive({
    req(plot_data_2())
    if (!input$same_radius) {
      radius <- axis_vars[axis_vars == input$radius_var]
      orig_imput_values <-
        zero_vars() %>%
        filter(var_code == radius) %>%
        gather(var, val, Original, Imputed) %>%
        pull(val)
    } else {
      orig_imput_values <- 1
      radius <- NULL
    }

    if (!is.null(radius) && all(orig_imput_values == 0)) radius <- NULL

    plot_scattered_countries(
      plot_data_2(),
      x_var = axis_vars[axis_vars == input$x_var],
      y_var = axis_vars[axis_vars == input$y_var],
      radius_var = radius,
      colour_var = group_vars[group_vars == input$colour_var]
    )
  })

```



Row {.tabset}
-------------------------------------

### Stage of Agricultural Development

If the plot does not appear, check the data summary table. It is possible, that there is no data for one or more variable and therefore, the plot cannot be built.
    
```{r}
fluidPage(
  fluidRow(
    renderUI(
      zero_message()
    ),
    renderPlot({
      p <- build_plot()
      print(p)
    }, width = "auto", height = 450)
  )
)

```

### Data availability and data export

```{r}
fluidPage(
  # fluidRow(
  #   tags$h3("Data availability")
  # ),
  # fluidRow(
  #   DT::renderDataTable( #
  #     DT::datatable(
  #       zero_vars() %>%
  #         select(-var_code),
  #       rownames = FALSE,
  #       options = list(
  #         pageLength = 25,
  #         dom = c("Brti")
  #       )
  #     )
  #   )
  # ),
  fluidRow(tags$h3("Data for export")),
  fluidRow(
    DT::renderDataTable( #
      DT::datatable(
        export_data(),
        rownames = FALSE,
        extensions = c("Buttons", "Scroller"),
        options = list(
          pageLength = 40,
          dom = c("Bfrtip"),
          scrollX = TRUE,
          # fixedColumns = list(leftColumns = 1),
          deferRender = TRUE,
          scrollY = 600,
          scroller = TRUE,
          buttons = list(
            list(extend = "copy", text = "Copy table"),
            list(extend = "excel", text = "Download table", title = "availability")
          )
        )
      )
    )
  )
)
```


### Clusters

```{r}
colours_calc
```




Data availability
=====================================  

Number of countries by variable and year. We do not show data for 1960-1990 because there is very little relevant data there. We also do extrapolation in order to fill the gaps, when data does not exist.

```{r}
availability <- 
  pd %>%
  group_by(Source, `Variable group`, var, flag, year) %>%
  count() %>%
  ungroup() %>%
  filter(year > 1995) %>%
  # mutate(val = str_c(flag, " ", n, "\n")) %>%
  # select(-n) %>%
  ungroup() %>%
  spread(flag, n) %>%
  mutate_at(vars(Imputed, Missing, Original), funs(ifelse(is.na(.), 0, .))) %>%
  mutate(status = str_c(
    ifelse(Original > 0, str_c("Original: ", Original, "\n"), ""),
    ifelse(Imputed > 0, str_c("Imputed: ", Imputed, "\n"), ""),
    ifelse(Missing > 0, str_c("Missing: ", Missing, "\n"), "")
  )) %>%
  select(-Imputed,-Missing,-Original) %>%
  spread(year, status) %>%
  arrange(Source, `Variable group`, var) %>%
  rename(Variable = var) %>%
  select(matches("\\D"), sort(names(.)[str_detect(names(.), "^(\\d{4})")], decreasing = T))

DT::renderDataTable( #
  DT::datatable(
    availability,
    rownames = FALSE,
    extensions = c("Buttons", "Scroller"),
    options = list(
      pageLength = 25,
      dom = c("Bfrtip"),
      scrollX = TRUE,
      # fixedColumns = list(leftColumns = 1),
      deferRender = TRUE,
      scrollY = 400,
      scroller = TRUE,
      buttons = list(
        list(extend = "copy", text = "Copy table"),
        list(extend = "excel", text = "Download table", title = "availability")
      )
    )
  )
)
```

